### 2020-11-18 latex/vlp.tex
Updated VLP according to new defn of energy gain

% TODO: Describe the oscillation and rotation motion of a pendulum in the
% qp-plane, and give a figure showing the orbiting motion for each of these.
% Then describe what the orbit should do under energy injection (spiral away /
% begin rotation outward) and energy dissipation (spiral in / rotate inward).

@@

\begin{thm}\label{thm:vlp-energy-stabilization}
   Define \(\theta := \arctan_2(p,q)\).
   Let \(L : \Sone \rightarrow [\underline{l},\overline{l}]\) be a
   \(C^2\) function of \(\theta\).
   A regular VNHC of the form 
   \(h(\mathbf{q},\mathbf{p}) = l - L(\theta)\) for the VLP injects energy 
   on the constraint manifold \(\Gamma \simeq \SxR\) if there
   exists \(l_\text{avg} \in [\underline{l},\overline{l}]\) such that 
   \begin{equation}\label{eqn:vlp-energy-gain-condition}
      \left(l_\text{avg} - L(\theta)\right)sin(2\theta) \geq 0 
      \text{ }\forall \theta \in \Sone
      ,
   \end{equation}
   with the property that the inequality is strict except at 
   \(\theta \in \{0, \frac{\pi}{2}, \frac{\pi}, \frac{3\pi}{2}\}\).  
   If instead 
   \[
    \left(l_\text{avg} - L(\theta)\right)sin(2\theta) \leq 0 
         \text{ }\forall \theta \in \Sone
         ,
   \]
   the VNHC dissipates energy on \(\Gamma\).
\end{thm}
\begin{proof}
We will first show that, under the given assumptions, the origin of the
constrained dynamics~\eqref{eqn:vlp-hamiltonian} is a repeller, and for that
we consider the negative time system obtained
from~\eqref{eqn:vlp-hamiltonian} by reversing the sign of the vector field:
\begin{equation}\label{eqn:negative-time-vlp}
   \begin{cases}
      \dot{q} = -\frac{p}{m L^2} \\
      \dot{p} = mgL\sin(q)
      . \\ 
   \end{cases}
\end{equation}
The state space of~\eqref{eqn:negative-time-vlp} is 
\(\SxR\), and the system has two equilibria,
\((q,p) = (0,0)\) and \((q,p) = (\pi,0)\). 
We need to show that the equilibrium \((q,p) = (0,0)\) is
asymptotically stable for~\eqref{eqn:negative-time-vlp}.

@@

% TODO: Add a remark:
% Our controller not only injects energy, but makes the lower equilibrium a
% repeller which means the "bad" set of x(0) does not contain a nbhd of (0,0),
% so anything near (0,0) will start to gain eenrgy aka ANY slight deviation from
% 0 will gain energy.

>>>>

@@

\begin{thm}\label{thm:vlp-energy-stabilization}
   Define \(\theta := \arctan_2(p,q)\).
   Let \(L : \Sone \rightarrow [\underline{l},\overline{l}]\) be a
   \(C^2\) function of \(\theta\).
   A regular VNHC of the form 
   \(h(\mathbf{q},\mathbf{p}) = l - L(\theta)\) for the VLP injects energy 
   on the constraint manifold \(\Gamma \simeq \SxR\) if there
   exists \(l_\text{avg} \in [\underline{l},\overline{l}]\) such that 
   \begin{equation}\label{eqn:vlp-energy-gain-condition}
      \left(l_\text{avg} - L(\theta)\right)sin(2\theta) \geq 0 
      \text{ }\forall \theta \in \Sone
      ,
   \end{equation}
   with the property that the inequality is strict except at 
   \(\theta \in \{0, \frac{\pi}{2}, \frac{\pi}, \frac{3\pi}{2}\}\).  
   If instead 
   \[
    \left(l_\text{avg} - L(\theta)\right)sin(2\theta) \leq 0 
         \text{ }\forall \theta \in \Sone
         ,
   \]
   the VNHC dissipates energy on \(\Gamma\).
\end{thm}
\begin{proof}
Since a system that gains energy per Definition \ref{defn:energy-gain}
cannot have stable equilibria,
we will first show that, under the given assumptions, the origin of the
constrained dynamics~\eqref{eqn:vlp-hamiltonian} is a repeller. 
For that we consider the negative time system obtained
from~\eqref{eqn:vlp-hamiltonian} by reversing the sign of the vector field:
\begin{equation}\label{eqn:negative-time-vlp}
   \begin{cases}
      \dot{q} = -\frac{p}{m L^2} \\
      \dot{p} = mgL\sin(q)
      . \\ 
   \end{cases}
\end{equation}
The state space of~\eqref{eqn:negative-time-vlp} is 
\(\SxR\), and the system has two equilibria,
\((q,p) = (0,0)\) and \((q,p) = (\pi,0)\). 
We need to show that the equilibrium \((q,p) = (0,0)\) is
asymptotically stable for~\eqref{eqn:negative-time-vlp}.

@@

% TODO: Add a remark:
% Our controller not only injects energy, but makes the lower equilibrium a
% repeller which means the "bad" set of x(0) does not contain a nbhd of (0,0),
% so anything near (0,0) will start to gain eenrgy aka ANY slight deviation from
% 0 will gain energy.
\begin{remark}
   Theorem \ref{thm:vlp-energy-stabilization} not only guarantees that our VNHC
   only injects energy, but that the lower equilibrium is a repeller. 
   If the VLP is initialized at the origin, any slight push along
   \(q\) will allow it to start gaining energy.
\end{remark}
